### 2021-04-11 dev_vignettes/Bushplus_strains.Rmd
testing

---
title: "Biodiversity, eco-evo feedback, and the stability of responses to environmental change"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(here)
library(deSolve)
library(rootSolve)

set.seed(13)

run_sims <- FALSE
```

# Introduction

This is relevant: https://www.biorxiv.org/content/10.1101/2021.03.22.436420v1

Aims:

* Investigate the effects of biodiversity within each of the three functional groups on responses to environmental change.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

We have added:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.
* Option to include biodiversity within functional groups.

# Methods

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Dynamic model

```{r}
source(here("functions/bushplus_dynamic_model.r"))
## contains various functions:
## bushplus_dynamic_model
## bushplus_dynamic_model_strains
```

## Defaults

### Default parameter values

To do:

* Make a function that can create parameter value sets, including for example controlling the number and types of strains within each functional group.

```{r}
default_parameter_values <- c(

  ## Maximum growth rate
  g_max_CB_0 = 0.05,
  g_max_CB_1 = 0.05,
  g_max_PB_0 = 0.07,
  g_max_PB_1 = 0.07,
  g_max_SB_0 = 0.10,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB_0 = 300,
  h_SR_CB_1 = 300,
  h_O_PB_0 = 100,
  h_O_PB_1 = 100,
  h_O_SB_0 = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  P_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB = 0,
  i_PB = 0,
  i_SB = 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5

  )
```

### Default initial state

To do:

* Make a function that can create initial conditions corresponding to the number of strains in the parameter values, and assign initial state values according to some rule (e.g. total abundance and even distribution of among strains).

Initial conditions favouring anoxic

```{r}
default_initial_state <- c(
  CB_0 = 5e1,
  CB_1 = 0,
  PB_0 = 1e7,
  PB_1 = 0,
  SB_0 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1)
```

### Default event

#### Default event definition

Used to add noise and to ensure minimum abundance.

Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

```{r}
source(here("functions/default_event_definition.r"))
## contains two functions:
## default_event_definition
## default_event_definition_strains -- this should work for any number of strains
```

#### Default event schedule

This is when an event happens

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimums

```{r}
## we only have one minimum abundance for each functional group
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB",
                                       "PB",
                                       "SB") 
```

### Default simulation duration

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 10
```

### Default oxygen diffusivity (a) forcing

Here we specify no forcing of a_0 (used unless set elsewhere).

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
default_log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(default_sim_duration),
                                            length=length(default_log10a_series))),
                                default_log10a_series))

```

## Simulation function

```{r}
source(here("functions/run_simulation.r"))
```

## Helper functions

```{r}
source(here("functions/plot_dynamics.r"))
source(here("functions/ss_expt_a_N.r"))
source(here("functions/get_nonlinearity.r"))
source(here("functions/get_hysteresis_range.r"))
```

## Code test

```{r}
dynamic_model = bushplus_dynamic_model_strains
event_definition = default_event_definition_strains
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_forcing = default_log10a_forcing
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing_func <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing_func, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
                           event_definition = default_event_definition_strains,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation()

plot_dynamics_strains(sim_res)

```

# Results: Reproducing previous results

Here we use the default parameter values, which includes two strains, but with same parameters. Results should be the same as with only one strain.

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(

  CB_0 = 5e1,
  CB_1 = 0,
  PB_0 = 1e7,
  PB_1 = 0,
  SB_0 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1

  )

sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
plot_dynamics_strains(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

## Initial conditions favouring anoxic, with noise, no forcing

Figure 2a, b in Bush et al, but with noise

```{r}
## add some noise
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          noise_sigma = noise_sigma_0.01)
plot_dynamics_strains(sim_res)
#ggsave("high_SB_PB.png")

noise_sigma_0.01 <- 0.00

```

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r}
noise_sigma_0.01 <- 0.00

init_state_favouring_oxic <- c(
  CB_0 = 1e7,
  CB_1 = 0,
  PB_0 = 1e5,
  PB_1 = 0,
  SB_0 = 1e5,
  SO = 500,
  SR = 50,
  O = 30,
  P = 10
  )

sim_res <- run_simulation(initial_state = init_state_favouring_oxic)

plot_dynamics_strains(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Initial conditions favouring oxic, with noise, no forcing

Figure 2c, d in Bush et al. but with noise.

```{r} 
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          noise_sigma = noise_sigma_0.01)

plot_dynamics_strains(sim_res)
##ggsave("high_CB.png")
noise_sigma_0.01 <- 0.00
```

# Results: two different CB strains

## Starting oxic steady state

```{r}
names(default_parameter_values)
two_CB_strains_parvals <- default_parameter_values
two_CB_strains_parvals["g_max_CB_1"] <- 0.0495
two_CB_strains_parvals["h_SR_CB_1"] <- 315
two_CB_strains_parvals

two_CB_init_state_favouring_oxic <- c(
  CB_0 = 0.75e7,
  CB_1 = 0.25e7,
  PB_0 = 1e5,
  PB_1 = 0,
  SB_0 = 1e5,
  SO = 500,
  SR = 50,
  O = 30,
  P = 10
  )

sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = two_CB_strains_parvals)

plot_dynamics_strains(sim_res)

```

## Forcing
```{r}

## simulation timing
simulation_duration <- 70000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -10, -10, -10, -10, -10, -10)

## run simulation with both CB strains having default growth rate
## and tolerance
sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = default_parameter_values,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)

plot_dynamics_strains(sim_res)
```

```{r}
## run simulation with CB strains having different growth rate
## and tolerance
sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = two_CB_strains_parvals,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)

plot_dynamics_strains(sim_res)

```

# TO DO:

* Check what happens when there are more strains.
* Be able to find the stable states.

# Results: Composition/diversity and nature of response

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 100
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

### Each composition

#### No organisms

```{r eval = run_sims}

initial_CBs <- 0
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_no_biology <- ss_by_a_N(expt)
saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
```

```{r}
ss_no_biology <- readRDS("data/ss_no_biology.RDS")
```

```{r}
ss_no_biology %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_no_biology %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_CB <- ss_by_a_N(expt)
saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
```

```{r}
ss_only_CB <- readRDS("data/ss_only_CB.RDS")
```

```{r}
tempx1 <- ss_only_CB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  geom_path(data = filter(tempx1, Species == "PB"),
            aes(y=0.2), size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only CB")

ggsave("figures/CB.png", width = 3, height = 2)
```

```{r}
ss_only_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 0
initial_SBs <- 10^seq(0, 8, length=grid_num_N)
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_SB <- ss_by_a_N(expt)

saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
```

```{r}
ss_only_SB <- readRDS("data/ss_only_SB.RDS")
```

```{r}
ss_only_SB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_PB <- ss_by_a_N(expt)

saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
```

```{r}
ss_only_PB <- readRDS("data/ss_only_PB.RDS")
```

```{r}
tempx1 <- ss_only_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size = 0.8) +
  theme_light() +
  ylim(0,10) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  geom_path(data = filter(tempx1, Species == "CB"),
            aes(y=0.2), size = 0.8)+
  geom_path(data = filter(tempx1, Species == "PB"),
            size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only PB")

ggsave("figures/PB.png", width = 3, height = 2)

```

```{r}
ss_only_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_PB

ss_only_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_CB

ss_only_SB_CB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:5) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_CB <- expt$N_PB

ss_only_CB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB, SB, and PB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB") 
ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### All response measures

```{r eval = run_sims}
all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
                    composition = rep("No organisms", grid_num_a*grid_num_N),
                    direction = c(rep("up", grid_num_a),
                                  rep("down", grid_num_a)))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB.RDS"),
                              composition = rep("CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB.RDS"),
                              composition = rep("SB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_PB.RDS"),
                              composition = rep("PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
                              composition = rep("SB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
                              composition = rep("SB-CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
                              composition = rep("CB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
                              composition = rep("CB-SB-PB",
                                                grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))

x_long <- all_ss %>%
  gather(key = Species, value = Quantity, 2:8)
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(composition, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(composition = fct_relevel(composition,
                                   "No organisms",
                                   "CB",
                                   "SB",
                                   "PB",
                                   "CB-PB",
                                   "SB-CB",
                                   "SB-PB",
                                   "CB-SB-PB"),
         num_func_groups = case_when(composition =="No organisms" ~ 0,
                                     composition == "CB" ~ 1,
                                     composition == "SB" ~ 1,
                                     composition == "PB" ~ 1,
                                     composition == "CB-PB" ~ 2,
                                     composition == "SB-CB" ~ 2,
                                     composition == "SB-PB" ~ 2,
                                     composition == "CB-SB-PB" ~ 3
                                     ),
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
saveRDS(resp_summary, "data/resp_summary.RDS")
```

```{r}
resp_summary <- readRDS("data/resp_summary.RDS")
resp_summary %>%
  ggplot() +
  geom_point(aes(x = composition, y = value)) +
  facet_grid(Species ~ resp_measure, scales = "free") +
  coord_flip()

```

### Nonlinearity

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(composition, Species, num_func_groups, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(composition, ave_nl)) +
  facet_wrap( ~ Species)

```

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = sum(value) / unique(num_func_groups))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, composition) %>%
  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(composition) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_mean_hyst)) +
  xlab("Functional groups present") +
  ylab("Amount of hysteresis") +
  ylim(0,2) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)

```

# Results: Varying strength of inhibition

## Introduction

Three parameters set the strength of inhibition (half-inhibition constants):

* h_SR_CB = 300,
* h_O_PB = 100,
* h_O_SB = 100,

These work by the inhibition function:

$\frac{1}{1 + \frac{X}{h_X}}$

```{r}
inhibition
```

Plot some examples:

```{r}
default_parameter_values["h_SR_CB"]
SR <- 10^seq(1.4, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
  ylab("Multiplicative effect\non CB growth rate")
```

```{r}
default_parameter_values["h_O_PB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
  ylab("Multiplicative effect\non PB growth rate")
```

```{r}
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

```{r}
proportion_change <- 2
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

## Experiment

Increasing the inhibition half saturation constants weakens the inhibition.

```{r}
ssfind_minimum_abundances <- default_minimum_abundances
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = NA,
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

grid_num_mult <- 20
grid_num_N <- 2
grid_num_a <- 100

a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5

expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    inhib_const_mult = inhib_const_mults,
                    a_O = a_Os)
```

```{r eval = run_sims}
ss_h1 <- ss_by_a_h_N(expt)

saveRDS(ss_h1, "data/ss_h1.RDS")
```

```{r}
ss_h1 <- readRDS("data/ss_h1.RDS")
```

## Results

### Close to default inhibition

```{r}
inhib_const_mults
focal_inhibition_multiplier <- inhib_const_mults[14]

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Double strength inhibition

```{r}
## YES! 0.5 does mean twice as strong inhibition
focal_inhibition_multiplier <- 0.5

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Weaker inhibition

```{r}
## YES! 2 does mean half as strong inhibition
focal_inhibition_multiplier <- 2^0.4

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Nonlinearity

```{r}
x_long <- ss_h1 %>%
  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
  gather(key = Species, value = Quantity, 3:9) 
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(inhib_const_mult, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
```

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(inhib_const_mult, Species, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_nl)) +
  facet_wrap( ~ Species)

```

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult,  resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, inhib_const_mult) %>%
  summarise(mean_hyst = mean(value))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(inhib_const_mult) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_mean_hyst)) +
  xlab("Organismal tolerance") +
  ylab("Amount of hysteresis") +
  theme_light() +
  ylim(0,2)
ggsave("figures/inhibition_system_hysteresis.png", width = 2.5, height = 2)

```

# Results: Temporal state switching

## Switch anoxic-oxic-anoxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_anoxic

## simulation timing
simulation_duration <- 5000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          log10a_series = lhl_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/lhl_with_defaults.RDS")
```

```{r}
lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
plot_dynamics(lhl_with_defaults)
```

```{r eval = run_sims}
lhl_with_defaults$result %>%
  filter(between(time, 1000, 73000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

## Switch oxic-anoxic-oxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_oxic

## simulation timing
simulation_duration <- 50000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/hlh_with_defaults.RDS")
```

```{r}
hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
plot_dynamics(hlh_with_defaults)
```

```{r eval = run_sims}
hlh_with_defaults$result %>%
  filter(between(time, 1000, 100000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

# Appendices

## runsteady function not finding correct values

```{r eval = FALSE}

init_state_anoxic_2 <- c(N_CB = 1e1,
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 300,
           SR = 300,
           O = 1e-5,
           P = 1e1)

RS <- runsteady(y = init_state_anoxic_2,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y
RS <- runsteady(y = init_state_favouring_oxic,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y

```

## Separatrix

Note that this section not working / updated since the code was made into functions

### Effect of N_CB and oxygen diffusability

```{r eval = FALSE}

noise_sigma <- 0.0

get_final_states_N_CB <- function(x) {

  #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_CB"] <- x["N_CB"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)

  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                    times = times,
                    func = micro_mod_var,
                    parms = parameters,
                    method = "radau",
                    events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 50000)

state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_CB and oxygen diffusability
grid_num <- 50
expt <- expand.grid(N_CB = 10^seq(0, 8, length=grid_num),
                    a_O = 10^seq(-3.5, -2, length=grid_num))

result <- apply(expt, 1, function(x) get_final_states_N_CB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_CB=expt$N_CB,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_CB>2e8, 'high', 'low'))

result <- result %>%
  mutate(N_CB = ifelse(N_CB < minimum_abundance, minimum_abundance, N_CB))

result$highlow <- as.factor(result$highlow)
saveRDS(result, "data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result <- readRDS("data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}

#plot N_CB / P_CB against a_0
p <- ggplot(result) + 
  geom_line(aes(x=log10(a_O), y=log10(N_CB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial cyanobacteria abundance')
p <- p + geom_point(data=result, aes(x=log10(a_O), y=log10(initial_N_CB), color=highlow)) + labs(color = "Final CB concentration")
plot(p)
#ggsave("initial conditions_1.png")
```

```{r eval = FALSE}
#show hysterisis effects for all the parameters
#plot oxygen against diffusability
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(O))))+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxygen concentration]')+
  geom_segment(aes(x=log10(result$a_O[839]), y=log10(result$O[839]), xend=log10(result$a_O[838]), yend=log10(result$O[838])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=log10(result$a_O[1761]), y=log10(result$O[1761]), xend=log10(result$a_O[1762]), yend=log10(result$O[1762])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=-4.25, y=-0.45, xend=-3.75, yend=0.1), arrow=arrow(), colour="blue", size=0.5)+
  geom_curve(aes(x = -3.75, y = 1.9, xend = -4.25, yend = 1.5), arrow=arrow(), colour='blue', size = 0.5, curvature = -0.2)
#ggsave("hysterisis_CB_1.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(P))))+
  ggtitle("Hysterisis effect of phoshate concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[phosphate concentration]')
#ggsave("hysterisis_CB_2.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SO))))+
  ggtitle("Hysterisis effect of oxidized sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxydized sulfur concentration]')
#ggsave("hysterisis_CB_3.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SR))))+
  ggtitle("Hysterisis effect of reduced sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[reduced sulfur] concentration]')
#ggsave("hysterisis_CB_4.png")
```

These graphs show an at least qualitative match to the result in the paper.

### Effect of PB/SB depending on oxygen diffusability

```{r eval = FALSE}
noise_sigma <- 0.00
get_final_states_N_PB <- function(x) { 
  #give in N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_PB"] <- x["var"]
  state["N_SB"] <- x["var"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)
  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                         times = times,
                         func = micro_mod_var,
                         parms = parameters,
                         method = "radau", events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 5000)
state <- c(N_CB = 5, 
           N_PB = NA,#this one is varied
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_PB and oxygen diffusivity
grid_num <- 50
expt <- expand.grid(var=10^seq(0, 7, length=grid_num),
                    a_O = 10^seq(-3, -2, length=grid_num))

result2 <- apply(expt, 1, function(x) get_final_states_N_PB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_PB=expt$var,
                                 initial_N_SB=expt$var,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_PB>1e8, 'high', 'low'))

result2 <- result2 %>%
  mutate(N_PB = ifelse(N_PB < minimum_abundance, minimum_abundance, N_PB))

result2$highlow <- as.factor(result2$highlow)
saveRDS(result2, "data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result2 <- readRDS("data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
#plot N_PB against a_0
p <- ggplot(result2) + 
  geom_line(aes(x=log10(a_O), y=log10(N_PB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial PS and SR bacteria concentration')
p <- p + geom_point(data=result2,
                    aes(x=log10(a_O), y=log10(initial_N_PB), color=highlow))
p <- p + labs(color = "Final PB and SB concentration")
plot(p)
#ggsave("initial conditions_2.png")
```

## EWS analyses

Evaluation of the switch: 
Get the functions: 
```{r eval = F}
source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
```

Get state shifts and calculate parameters: 
```{r eval = F}
x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)

#first switch
name_1='anoxic_oxic 1'
trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
out_df_1_N_CB <- data.frame(trans1$N_CB)
colnames(out_df_1_N_CB) <- c("N_CB")
out_df_1_N_PB <- data.frame(trans1$N_PB)
colnames(out_df_1_N_PB) <- c("N_PB")
out_df_1_N_SB <- data.frame(trans1$N_SB)
colnames(out_df_1_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
  ggtitle("Anoxic_oxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 1.png")

#second switch
name_2='oxic_anoxic 1'
trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
n_23 <- round(length(trans2$time)*x3/100) 
t_23 <- length(trans2$time) - n_23 + 1
n_25 <- round(length(trans2$time)*x5/100) 
t_25 <- length(trans2$time) - n_25 + 1
out_df_2_N_CB <- data.frame(trans2$N_CB)
colnames(out_df_2_N_CB) <- c("N_CB")
out_df_2_N_PB <- data.frame(trans2$N_PB)
colnames(out_df_2_N_PB) <- c("N_PB")
out_df_2_N_SB <- data.frame(trans2$N_SB)
colnames(out_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
  ggtitle("Oxic_anoxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 1.png")

#third switch
name_3='oxic_anoxic 2'
trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
n_33 <- round(length(trans3$time)*x3/100) 
t_33 <- length(trans3$time) - n_33 + 1
n_35 <- round(length(trans3$time)*x5/100) 
t_35 <- length(trans3$time) - n_35 + 1
out2_df_1_N_CB <- data.frame(trans3$N_CB)
colnames(out2_df_1_N_CB) <- c("N_CB")
out2_df_1_N_SB <- data.frame(trans3$N_SB)
colnames(out2_df_1_N_SB) <- c("N_SB")
out2_df_1_N_PB <- data.frame(trans3$N_PB)
colnames(out2_df_1_N_PB) <- c("N_PB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
  ggtitle("Oxic_anoxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 2.png")

#fourth switch
name_4='anoxic_oxic 2'
trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
n_43 <- round(length(trans4$time)*x3/100) 
t_43 <- length(trans4$time) - n_43 + 1
n_45 <- round(length(trans4$time)*x5/100) 
t_45 <- length(trans4$time) - n_45 + 1
out2_df_2_N_CB <- data.frame(trans4$N_CB)
colnames(out2_df_2_N_CB) <- c("N_CB")
out2_df_2_N_PB <- data.frame(trans4$N_PB)
colnames(out2_df_2_N_PB) <- c("N_PB")
out2_df_2_N_SB <- data.frame(trans4$N_SB)
colnames(out2_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
  ggtitle("Anoxic_oxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 2.png")
```

Rolling window analysis
```{r eval = F}
#analysis for 1st
rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)

```

Metric based indicators from the earlywarnings package
```{r eval = F}
#analysis for 1st
earlywarnings_plot(out_df_1_N_CB, name_1, x3)
earlywarnings_plot(out_df_1_N_PB, name_1, x3)
earlywarnings_plot(out_df_1_N_SB, name_1, x3)

#analysis for 2nd
earlywarnings_plot(out_df_2_N_CB, name_2, x3)
earlywarnings_plot(out_df_2_N_PB, name_2, x3)
earlywarnings_plot(out_df_2_N_SB, name_2, x3)

#analysis for 3rd
earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
earlywarnings_plot(out2_df_1_N_SB, name_3, x3)

#analysis for 4th
earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
```

Conditional Heteroskedasticity 
```{r eval = F}
#analysis for 1st
ch <- condhet_analysis(out_df_1_N_CB, name_1)
ch <- condhet_analysis(out_df_1_N_PB, name_1)
ch <- condhet_analysis(out_df_1_N_SB, name_1)

#analysis for 2nd
ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
ch2 <- condhet_analysis(out_df_2_N_SB, name_2)

#analysis for 3rd
ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)

#analysis for 4th
ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
```

Detrending fluctuation analysis 
```{r eval = F}
#analysis for 1st
dfa_analysis(trans1, 2, n_15, t_15, name_1)
dfa_analysis(trans1, 3, n_15, t_15, name_1)
dfa_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
dfa_analysis(trans2, 2, n_25, t_25, name_2)
dfa_analysis(trans2, 3, n_25, t_25, name_2)
dfa_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
dfa_analysis(trans3, 2, n_35, t_35, name_3)
dfa_analysis(trans3, 3, n_35, t_35, name_3)
dfa_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
dfa_analysis(trans4, 2, n_45, t_45, name_4)
dfa_analysis(trans4, 3, n_45, t_45, name_4)
dfa_analysis(trans4, 4, n_45, t_45, name_4)
```

Drift Diffusion Jump Nonparametrics Early Warning Signals:
```{r eval = F}
#shift 1
ddj11<- ddj_model(out_df_1_N_CB)
plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
plot_ddj_2(out_df_1_N_CB, ddj11, name_1)

ddj12<- ddj_model(out_df_1_N_PB)
plot_ddj_1(out_df_1_N_PB, ddj12, name_1)
plot_ddj_2(out_df_1_N_PB, ddj12, name_1)

ddj13<- ddj_model(out_df_1_N_SB)
plot_ddj_1(out_df_1_N_SB, ddj13, name_1)
plot_ddj_2(out_df_1_N_SB, ddj13, name_1)

#shift 2
ddj21<- ddj_model(out_df_2_N_CB)
plot_ddj_1(out_df_2_N_CB, ddj21, name_2)
plot_ddj_2(out_df_2_N_CB, ddj21, name_2)

ddj22<- ddj_model(out_df_2_N_PB)
plot_ddj_1(out_df_2_N_PB, ddj22, name_2)
plot_ddj_2(out_df_2_N_PB, ddj22, name_2)

ddj23<- ddj_model(out_df_2_N_SB)
plot_ddj_1(out_df_2_N_SB, ddj23, name_2)
plot_ddj_2(out_df_2_N_SB, ddj23, name_2)

#shift 3
ddj31<- ddj_model(out2_df_1_N_CB)
plot_ddj_1(out2_df_1_N_CB, ddj31, name_3)
plot_ddj_2(out2_df_1_N_CB, ddj31, name_3)

ddj32<- ddj_model(out2_df_1_N_PB)
plot_ddj_1(out2_df_1_N_PB, ddj32, name_3)
plot_ddj_2(out2_df_1_N_PB, ddj32, name_3)

ddj33<- ddj_model(out2_df_1_N_SB)
plot_ddj_1(out2_df_1_N_SB, ddj33, name_3)
plot_ddj_2(out2_df_1_N_SB, ddj33, name_3)

#shift 4
ddj41<- ddj_model(out2_df_2_N_CB)
plot_ddj_1(out2_df_2_N_CB, ddj41, name_4)
plot_ddj_2(out2_df_2_N_CB, ddj41, name_4)

ddj42<- ddj_model(out2_df_2_N_PB)
plot_ddj_1(out2_df_2_N_PB, ddj42, name_4)
plot_ddj_2(out2_df_2_N_PB, ddj42, name_4)

ddj43<- ddj_model(out2_df_2_N_SB)
plot_ddj_1(out2_df_2_N_SB, ddj43, name_4)
plot_ddj_2(out2_df_2_N_SB, ddj43, name_4)
```