### 2021-04-13 various_stuff/dev_vignettes/Bush_basic_reproduction_new.Rmd
lots

---
title: "Reproducing some results of Bush et al 2017"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(deSolve)
library(rootSolve)
library(here)
library(microxanox)

## Set random number generator seed
## Is redundant if there are no stochastic components in the model
set.seed(13)

run_steadystate_sims <- FALSE
```

# Introduction

The aims here are:

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
* Add an investigation of how changes in tolerances alter the response to environmental change.

We also add:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.

Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups. That is, there is only one parameter set per functional group.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

# Methods

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Modelling framework

The framework used for running a simulation, and for varying aspects of a simulation run is:

* Define the default simulation conditions.
* Use these defaults unless other conditions are specified.
* The function call `run_simulation()` would therefore run a simulation with default conditions.
* Whereas, for example, the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values.

### Dynamic model

The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model_strains()`. This is an augemented version that can handle multiple strains within each of the three functional groups. The basic version that has only one strain per functional group is `bushplus_dynamic_model()`.

### Default parameter values

```{r}
default_parameter_values <- c(

  ## maximum specific growth rates
  g_max_CB = 0.05,
  g_max_PB = 0.07,
  g_max_SB = 0.1,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB = 300,
  h_O_PB = 100,
  h_O_SB = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  Pr_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB <- 0,
  i_PB <- 0,
  i_SB <- 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5
  )
```

### Default event

#### Default event definition

Used to add noise and to ensure minimum abundance. Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

The event definition is made in the form of a function. The default definition is in the function `default_event_definition()`

#### Default event schedule

At the moment this is specified by giving the time interval between events:

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise added to substrate concentrations:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimum abundances:

```{r}
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
```

### Default simulation duration

Default to the following duration and interval between samples:

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 1
```

### Default oxygen diffusivity (a) forcing

Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
```

### Default initial state

Specify the default set of initial conditions (here favouring anoxic organisms).

```{r}
default_initial_state <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)
```

## Code test

```{r}
dynamic_model = bushplus_dynamic_model
event_definition = default_event_definition
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_series = default_log10a_series
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(times),
                                            length=length(log10a_series))),
                                log10a_series))

log10a_forcing <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

bushplus_dynamic_model(1, initial_state, default_parameter_values)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
                           event_definition = default_event_definition,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)

plot_dynamics(sim_res)

```

# Results

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)

parameters <- default_parameter_values
state <- init_state_favouring_anoxic
bushplus_dynamic_model(100, state, parameters)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

Population dynamics seem as in the paper.
Resource dynamics fluctuate more than in the paper.

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r} 
init_state_favouring_oxic <- c(N_CB = 1e7,
           N_PB = 1e5,
           N_SB = 1e5,
           SO = 500,
           SR = 50,
           O = 30,
           P = 10)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_oxic)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Stable states

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 2
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

```{r eval = run_steadystate_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
```

```{r}
#ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB")  %>%
  print()
##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
p1 <- ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
print(p1)
```

# Repeat the results above using the functions that allow simulation of multiple strains per functional group

## Test with one strain per functional group

Constructing parameter set.

```{r}
CB_strain_params <- data.frame(strain_name = c("CB_1"),
                       g_max_CB = c(0.05),
                       k_CB_P = c(0.2),
                       h_SR_CB = c(300),
                       y_P_CB = c(1.67e8),
                       Pr_CB = c(6e-9),
                       m_CB = c(0.02),
                       i_CB = c(0))

PB_strain_params <- data.frame(strain_name = c("PB_1"),
                       g_max_PB = c(0.07),
                       k_PB_SR = 10,
                       k_PB_P = 0.5,
                       h_O_PB = 100,
                       y_SR_PB = 1.25e7,
                       y_P_PB = 1.67e8,
                       m_PB = 0.028,
                       i_PB = 0)

SB_strain_params <- data.frame(strain_name = c("SB_1"),
                       g_max_SB = c(0.1),
                       k_SB_SO = 5,
                       k_SB_P = 0.5,
                       h_O_SB = 100,
                       y_SO_SB = 3.33e7,  
                       y_P_SB = 1.67e8,
                       m_SB = 0.04,
                       i_SB = 0)

all_other_params <- c(

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## oxidisation rate of reduced sulphur
  c = 4e-5

)

one_strain_params <- construct_parameter_set(CB_strain_params,
                                    PB_strain_params,
                                    SB_strain_params,
                                    all_other_params)

one_strain_initial_state <- c(
  CB_1 = 5e1,
  PB_1 = 1e7,
  SB_1 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1
  )

simulation_result <- run_simulation(dynamic_model =
                                      bushplus_dynamic_model_strains,
                                    parameter_values = one_strain_params,
                                    initial_state = one_strain_initial_state)

plot_dynamics_strains(simulation_result)

```

### Test with N strains per functional group

```{r}
num_CB_strains <- 3
num_PB_strains <- 3
num_SB_strains <- 3

CB_strain_params <- data.frame(strain_name = paste0("CB_", 1:num_CB_strains),
                       g_max_CB = rep(0.05, num_CB_strains),
                       k_CB_P = rep(0.2, num_CB_strains),
                       h_SR_CB = rep(300, num_CB_strains),
                       y_P_CB = rep(1.67e8, num_CB_strains),
                       Pr_CB = rep(6e-9, num_CB_strains),
                       m_CB = rep(0.02, num_CB_strains),
                       i_CB = rep(0, num_CB_strains))

PB_strain_params <- data.frame(strain_name = paste0("PB_", 1:num_PB_strains),
                       g_max_PB = rep(0.07, num_PB_strains),
                       k_PB_SR = rep(10, num_PB_strains),
                       k_PB_P = rep(0.5, num_PB_strains),
                       h_O_PB = rep(100, num_PB_strains),
                       y_SR_PB = rep(1.25e7, num_PB_strains),
                       y_P_PB = rep(1.67e8, num_PB_strains),
                       m_PB = rep(0.028, num_PB_strains),
                       i_PB = rep(0, num_PB_strains))

SB_strain_params <- data.frame(strain_name = paste0("SB_", 1:num_SB_strains),
                       g_max_SB = rep(0.1, num_SB_strains),
                       k_SB_SO = rep(5, num_SB_strains),
                       k_SB_P = rep(0.5, num_SB_strains),
                       h_O_SB = rep(100, num_SB_strains),
                       y_SO_SB = rep(3.33e7, num_SB_strains),  
                       y_P_SB = rep(1.67e8, num_SB_strains),
                       m_SB = rep(0.04, num_SB_strains),
                       i_SB = rep(0, num_SB_strains))

multi_strain_params <- construct_parameter_set(CB_strain_params,
                                    PB_strain_params,
                                    SB_strain_params,
                                    all_other_parms)

multi_strain_initial_state <- c(
  CB_1 = 5e1,
  CB_2 = 0,
  CB_3 = 0,
  PB_1 = 1e7,
  PB_2 = 1e6,
  PB_3 = 1e5,
  SB_1 = 1e7,
  SB_2 = 1e6,
  SB_3 = 1e5,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1)

simulation_result <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
               parameter_values = multi_strain_params,
               initial_state = multi_strain_initial_state)
plot_dynamics_strains(simulation_result)

```