### 2021-04-14 various_stuff/EWS functions/EWS analyses.Rmd
reorganisation of vignettes
now need to get them working!

---
title: "EWS analysis"
author: "Owen Petchey & Jennifer Probst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

## EWS analyses

Evaluation of the switch: 
Get the functions: 
```{r eval = F}
source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
```

Get state shifts and calculate parameters: 
```{r eval = F}
x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)

#first switch
name_1='anoxic_oxic 1'
trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
out_df_1_N_CB <- data.frame(trans1$N_CB)
colnames(out_df_1_N_CB) <- c("N_CB")
out_df_1_N_PB <- data.frame(trans1$N_PB)
colnames(out_df_1_N_PB) <- c("N_PB")
out_df_1_N_SB <- data.frame(trans1$N_SB)
colnames(out_df_1_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
  ggtitle("Anoxic_oxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 1.png")

#second switch
name_2='oxic_anoxic 1'
trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
n_23 <- round(length(trans2$time)*x3/100) 
t_23 <- length(trans2$time) - n_23 + 1
n_25 <- round(length(trans2$time)*x5/100) 
t_25 <- length(trans2$time) - n_25 + 1
out_df_2_N_CB <- data.frame(trans2$N_CB)
colnames(out_df_2_N_CB) <- c("N_CB")
out_df_2_N_PB <- data.frame(trans2$N_PB)
colnames(out_df_2_N_PB) <- c("N_PB")
out_df_2_N_SB <- data.frame(trans2$N_SB)
colnames(out_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
  ggtitle("Oxic_anoxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 1.png")

#third switch
name_3='oxic_anoxic 2'
trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
n_33 <- round(length(trans3$time)*x3/100) 
t_33 <- length(trans3$time) - n_33 + 1
n_35 <- round(length(trans3$time)*x5/100) 
t_35 <- length(trans3$time) - n_35 + 1
out2_df_1_N_CB <- data.frame(trans3$N_CB)
colnames(out2_df_1_N_CB) <- c("N_CB")
out2_df_1_N_SB <- data.frame(trans3$N_SB)
colnames(out2_df_1_N_SB) <- c("N_SB")
out2_df_1_N_PB <- data.frame(trans3$N_PB)
colnames(out2_df_1_N_PB) <- c("N_PB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
  ggtitle("Oxic_anoxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 2.png")

#fourth switch
name_4='anoxic_oxic 2'
trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
n_43 <- round(length(trans4$time)*x3/100) 
t_43 <- length(trans4$time) - n_43 + 1
n_45 <- round(length(trans4$time)*x5/100) 
t_45 <- length(trans4$time) - n_45 + 1
out2_df_2_N_CB <- data.frame(trans4$N_CB)
colnames(out2_df_2_N_CB) <- c("N_CB")
out2_df_2_N_PB <- data.frame(trans4$N_PB)
colnames(out2_df_2_N_PB) <- c("N_PB")
out2_df_2_N_SB <- data.frame(trans4$N_SB)
colnames(out2_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
  ggtitle("Anoxic_oxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 2.png")
```

Rolling window analysis
```{r eval = F}
#analysis for 1st
rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)

```

Metric based indicators from the earlywarnings package
```{r eval = F}
#analysis for 1st
earlywarnings_plot(out_df_1_N_CB, name_1, x3)
earlywarnings_plot(out_df_1_N_PB, name_1, x3)
earlywarnings_plot(out_df_1_N_SB, name_1, x3)

#analysis for 2nd
earlywarnings_plot(out_df_2_N_CB, name_2, x3)
earlywarnings_plot(out_df_2_N_PB, name_2, x3)
earlywarnings_plot(out_df_2_N_SB, name_2, x3)

#analysis for 3rd
earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
earlywarnings_plot(out2_df_1_N_SB, name_3, x3)

#analysis for 4th
earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
```

Conditional Heteroskedasticity 
```{r eval = F}
#analysis for 1st
ch <- condhet_analysis(out_df_1_N_CB, name_1)
ch <- condhet_analysis(out_df_1_N_PB, name_1)
ch <- condhet_analysis(out_df_1_N_SB, name_1)

#analysis for 2nd
ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
ch2 <- condhet_analysis(out_df_2_N_SB, name_2)

#analysis for 3rd
ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)

#analysis for 4th
ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
```

Detrending fluctuation analysis 
```{r eval = F}
#analysis for 1st
dfa_analysis(trans1, 2, n_15, t_15, name_1)
dfa_analysis(trans1, 3, n_15, t_15, name_1)
dfa_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
dfa_analysis(trans2, 2, n_25, t_25, name_2)
dfa_analysis(trans2, 3, n_25, t_25, name_2)
dfa_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
dfa_analysis(trans3, 2, n_35, t_35, name_3)
dfa_analysis(trans3, 3, n_35, t_35, name_3)
dfa_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
dfa_analysis(trans4, 2, n_45, t_45, name_4)
dfa_analysis(trans4, 3, n_45, t_45, name_4)
dfa_analysis(trans4, 4, n_45, t_45, name_4)
```

Drift Diffusion Jump Nonparametrics Early Warning Signals:
```{r eval = F}
#shift 1
ddj11<- ddj_model(out_df_1_N_CB)
plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
plot_ddj_2(out_df_1_N_CB, ddj11, name_1)

ddj12<- ddj_model(out_df_1_N_PB)
plot_ddj_1(out_df_1_N_PB, ddj12, name_1)
plot_ddj_2(out_df_1_N_PB, ddj12, name_1)

ddj13<- ddj_model(out_df_1_N_SB)
plot_ddj_1(out_df_1_N_SB, ddj13, name_1)
plot_ddj_2(out_df_1_N_SB, ddj13, name_1)

#shift 2
ddj21<- ddj_model(out_df_2_N_CB)
plot_ddj_1(out_df_2_N_CB, ddj21, name_2)
plot_ddj_2(out_df_2_N_CB, ddj21, name_2)

ddj22<- ddj_model(out_df_2_N_PB)
plot_ddj_1(out_df_2_N_PB, ddj22, name_2)
plot_ddj_2(out_df_2_N_PB, ddj22, name_2)

ddj23<- ddj_model(out_df_2_N_SB)
plot_ddj_1(out_df_2_N_SB, ddj23, name_2)
plot_ddj_2(out_df_2_N_SB, ddj23, name_2)

#shift 3
ddj31<- ddj_model(out2_df_1_N_CB)
plot_ddj_1(out2_df_1_N_CB, ddj31, name_3)
plot_ddj_2(out2_df_1_N_CB, ddj31, name_3)

ddj32<- ddj_model(out2_df_1_N_PB)
plot_ddj_1(out2_df_1_N_PB, ddj32, name_3)
plot_ddj_2(out2_df_1_N_PB, ddj32, name_3)

ddj33<- ddj_model(out2_df_1_N_SB)
plot_ddj_1(out2_df_1_N_SB, ddj33, name_3)
plot_ddj_2(out2_df_1_N_SB, ddj33, name_3)

#shift 4
ddj41<- ddj_model(out2_df_2_N_CB)
plot_ddj_1(out2_df_2_N_CB, ddj41, name_4)
plot_ddj_2(out2_df_2_N_CB, ddj41, name_4)

ddj42<- ddj_model(out2_df_2_N_PB)
plot_ddj_1(out2_df_2_N_PB, ddj42, name_4)
plot_ddj_2(out2_df_2_N_PB, ddj42, name_4)

ddj43<- ddj_model(out2_df_2_N_SB)
plot_ddj_1(out2_df_2_N_SB, ddj43, name_4)
plot_ddj_2(out2_df_2_N_SB, ddj43, name_4)
```