### 2021-04-14 various_stuff/dev_vignettes/partial reproduction Bushetal2017.Rmd

---
title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(deSolve)
library(rootSolve)
library(here)
library(microxanox)

## Set random number generator seed
## Is redundant if there are no stochastic components in the model
set.seed(13)

run_steadystate_sims <- FALSE
```

# Introduction

The aims here are:

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).

Please see the User manual vignette for further introduction and general methods.

# Methods

### Default parameter values

Only one set of parameter values is used throughout. It gives the possibility for alternate stable states.

```{r}
default_parameter_values <- c(

  ## maximum specific growth rates
  g_max_CB = 0.05,
  g_max_PB = 0.07,
  g_max_SB = 0.1,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB = 300,
  h_O_PB = 100,
  h_O_SB = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  Pr_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB <- 0,
  i_PB <- 0,
  i_SB <- 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5
  )
```

### Default initial state

Specify the default set of initial conditions (here favouring anoxic organisms).

```{r}
default_initial_state <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)
```

### Default simulation duration

Default to the following duration and interval between samples:

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 1
```

### Default oxygen diffusivity (a) forcing

Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
```

### Default event

#### Default event definition

Default is to ensure abundances are not below minimum, and to add the amount of noise specified below (none in this vignette).

These default are in the definition function `default_event_definition()`

#### Default event schedule

Here set to an event every 10 time steps.

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise added to substrate concentrations:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimum abundances:

```{r}
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
```

## Code test

Remove this section when the rest definitely works.

```{r eval = FALSE}
dynamic_model = bushplus_dynamic_model
event_definition = default_event_definition
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_series = default_log10a_series
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(times),
                                            length=length(log10a_series))),
                                log10a_series))

log10a_forcing <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

bushplus_dynamic_model(1, initial_state, default_parameter_values)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
                           event_definition = default_event_definition,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)

plot_dynamics(sim_res)

```

# Results

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)

parameters <- default_parameter_values
state <- init_state_favouring_anoxic
bushplus_dynamic_model(100, state, parameters)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

Population dynamics seem as in the paper.
Resource dynamics fluctuate more than in the paper.

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r} 
init_state_favouring_oxic <- c(N_CB = 1e7,
           N_PB = 1e5,
           N_SB = 1e5,
           SO = 500,
           SR = 50,
           O = 30,
           P = 10)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_oxic)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Stable states

```{r eval = run_steadystate_sims}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 2
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
```

```{r}
load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB")  %>%
  print()
##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
p1 <- ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
print(p1)
```

## Temporal state switching

### Anoxic-oxic-anoxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_anoxic

## simulation timing
simulation_duration <- 5000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          log10a_series = lhl_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/lhl_with_defaults.RDS")
```

```{r}
lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
plot_dynamics(lhl_with_defaults)
```

```{r eval = run_sims}
lhl_with_defaults$result %>%
  filter(between(time, 1000, 73000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

### Oxic-anoxic-oxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_oxic

## simulation timing
simulation_duration <- 50000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/hlh_with_defaults.RDS")
```

```{r}
hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
plot_dynamics(hlh_with_defaults)
```

```{r eval = run_sims}
hlh_with_defaults$result %>%
  filter(between(time, 1000, 100000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

## Composition and nature of response

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 100
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

### Each composition

#### No organisms

```{r eval = run_sims}

initial_CBs <- 0
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_no_biology <- ss_by_a_N(expt)
saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
```

```{r}
ss_no_biology <- readRDS("data/ss_no_biology.RDS")
```

```{r}
ss_no_biology %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_no_biology %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_CB <- ss_by_a_N(expt)
saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
```

```{r}
ss_only_CB <- readRDS("data/ss_only_CB.RDS")
```

```{r}
tempx1 <- ss_only_CB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  geom_path(data = filter(tempx1, Species == "PB"),
            aes(y=0.2), size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only CB")

ggsave("figures/CB.png", width = 3, height = 2)
```

```{r}
ss_only_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 0
initial_SBs <- 10^seq(0, 8, length=grid_num_N)
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_SB <- ss_by_a_N(expt)

saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
```

```{r}
ss_only_SB <- readRDS("data/ss_only_SB.RDS")
```

```{r}
ss_only_SB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_PB <- ss_by_a_N(expt)

saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
```

```{r}
ss_only_PB <- readRDS("data/ss_only_PB.RDS")
```

```{r}
tempx1 <- ss_only_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size = 0.8) +
  theme_light() +
  ylim(0,10) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  geom_path(data = filter(tempx1, Species == "CB"),
            aes(y=0.2), size = 0.8)+
  geom_path(data = filter(tempx1, Species == "PB"),
            size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only PB")

ggsave("figures/PB.png", width = 3, height = 2)

```

```{r}
ss_only_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_PB

ss_only_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_CB

ss_only_SB_CB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:5) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_CB <- expt$N_PB

ss_only_CB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB, SB, and PB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB") 
ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### All response measures

```{r eval = run_sims}
all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
                    composition = rep("No organisms", grid_num_a*grid_num_N),
                    direction = c(rep("up", grid_num_a),
                                  rep("down", grid_num_a)))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB.RDS"),
                              composition = rep("CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB.RDS"),
                              composition = rep("SB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_PB.RDS"),
                              composition = rep("PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
                              composition = rep("SB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
                              composition = rep("SB-CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
                              composition = rep("CB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
                              composition = rep("CB-SB-PB",
                                                grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))

x_long <- all_ss %>%
  gather(key = Species, value = Quantity, 2:8)
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(composition, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(composition = fct_relevel(composition,
                                   "No organisms",
                                   "CB",
                                   "SB",
                                   "PB",
                                   "CB-PB",
                                   "SB-CB",
                                   "SB-PB",
                                   "CB-SB-PB"),
         num_func_groups = case_when(composition =="No organisms" ~ 0,
                                     composition == "CB" ~ 1,
                                     composition == "SB" ~ 1,
                                     composition == "PB" ~ 1,
                                     composition == "CB-PB" ~ 2,
                                     composition == "SB-CB" ~ 2,
                                     composition == "SB-PB" ~ 2,
                                     composition == "CB-SB-PB" ~ 3
                                     ),
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
saveRDS(resp_summary, "data/resp_summary.RDS")
```

```{r}
resp_summary <- readRDS("data/resp_summary.RDS")
resp_summary %>%
  ggplot() +
  geom_point(aes(x = composition, y = value)) +
  facet_grid(Species ~ resp_measure, scales = "free") +
  coord_flip()

```

### Nonlinearity

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(composition, Species, num_func_groups, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(composition, ave_nl)) +
  facet_wrap( ~ Species)

```

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = sum(value) / unique(num_func_groups))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, composition) %>%
  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(composition) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_mean_hyst)) +
  xlab("Functional groups present") +
  ylab("Amount of hysteresis") +
  ylim(0,2) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)

```