### 2022-06-16 inst/manuscript/manuscript_microxanox.tex
Changes Bibliography and re-compiled.

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{SoftwareX} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

%% \documentclass[preprint,12pt, a4paper]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

\usepackage{float}
\restylefloat{table}

\journal{SoftwareX}

%\usepackage{lineno}
%\linenumbers
%\usepackage{setspace}
%\doublespacing
%\usepackage{fancyhdr}
%\fancyhead{}
%\fancyfoot{}
%\fancyhead[CO,CE]{Microxanox R package}
%\pagestyle{fancy}
%

\begin{document}

\begin{frontmatter}

  \title{Microxanox: an R package for simulating an aquatic
\(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
    \author[University of Zürich]{Rainer M Krug%
  %
  \fnref{Corresponding Author}}
   \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
    \author[University of Zürich]{Owen L. Petchey}
   \ead{Owen.Petchey@ieu.uzh.ch} 
      \affiliation[University of Zürich]{Department of Evolutionary
Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}
    \fntext[2]{Equal contribution}

  \begin{abstract}
  \(Microxanox\) is an R package to simulate a three functional group
  ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
  oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
  differential equations. Simulations can be run individually or over a
  parameter range, to find stable states. The model can be implemented
  with different numbers of species per functional group. The package is
  constructed in such a way that the results contain the input parameter
  used, so that a saved results can be loaded again and the simulation
  be repeated. Furthermore, the package framework and code should serve
  as a useful starting point for making simulation models of other types
  of ecosystem.
  \end{abstract}
    \begin{keyword}
    reproducibility \sep regime shift \sep stable state \sep 
    ordinary differential equations
  \end{keyword}

 \end{frontmatter}

\pagebreak

\hypertarget{required-metadata}{%
\section{Required Metadata}\label{required-metadata}}

\hypertarget{current-code-version}{%
\subsection{Current code version}\label{current-code-version}}

Ancillary data table required for subversion of the codebase.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Owen.Petchey@ieu.uzh \\
\bottomrule
\end{longtable}

\hypertarget{motivation-and-significance}{%
\section{Motivation and
significance}\label{motivation-and-significance}}

Mathematical models play a key role in the development of understanding
about how ecosystems work and how they respond to environmental changes
\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
important for developing hypotheses to test in empirical studies. One
area of ecology in which models have played a influential role is how
ecosystems respond to gradual change in an environmental driver
\citep{Scheffer2001}. An environmental driver is a environmental
condition that affects an ecosystem, but is assumed to not be affected
by the ecosystem, such as the rate of nutrient input into a lake.

It is conceivable that an ecosystem state, such as the total biomass of
a particular type of bacteria, may remain unchanged when an
environmental driver changes. It is also possible that the ecosystem
state changes gradually. It is also possible that the ecosystem state
changes abruptly to a new state that is difficult to recover from
\citep{Scheffer2001}. This possibility for abrupt, perhaps catastrophic
changes that are difficult to reverse causes considerable concern
\citep{Collins2021, Northrop2021, Vandermeer2019}.

An example where a gradual change of an environmental variable causes an
abrupt change of the system is the switch from an aerobic (oxygen is
available for metabolism) to anaerobic (oxygen generally unavailable)
state in a microbial ecosystem. This system has been investigated by
\citet{Bush2017} in a simulation study of a mathematical model. Three
types of microbes occur in the model: cyanobacteria (CB) dominating the
oxic state, and two types of sulfur bacteria that dominate the anoxic
state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria
(PB)). The model shows that gradual change in the rate at which oxygen
could diffuse into the ecosystem (termed the oxygen diffusitivity) could
cause catastrophic changes in the ecosystem state that would be
difficult to reverse.

One feature of the study by \citet{Bush2017} was limited biodiversity.
Specifically, there was no biodiversity within each of the three types
of bacteria. This leaves open the question of if and how biodiversity
within these types (i.e.~functional groups) of bacteria affects the
ecosystem response to environmental change. This limitation is not
specific to the study of \citet{Bush2017}. There are few if any studies
of the effects of biodiversity on abrupt transitions between ecosystem
states.

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change \citet{Limberger2022}, and to base our work on the
work and model of \citet{Bush2017}. It was with this goal in mind that
we developed the \emph{microxanox} package \citep{Krug2022}. The first
stage of development was to write code from scratch (as there was no
available code to start from) and to confirm that this new
implementation would reproduce the previously published results. The
resulting reproduction is available as one of the package vignettes:
\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
Partial reproduction of Bush et al}.

The second stage was to add functionality that would be necessary to
answer our research question. Most importantly, we made it possible to
have multiple species of bacteria within each of the three functional
groups, for the multiple species to differ in their characteristics, and
to vary the number of species and amount of variability among them. We
also added functionality that allowed: temporally varying environmental
conditions, addition of random noise to state variables, and
immigration. In addition to the model itself, the package includes some
functions to analyse the results as well as to visualize the results to
provide a starting point for customized visualizations based on own
requirements. The basic and additional functionality is described in the
package
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

\hypertarget{software-description}{%
\section{Software description}\label{software-description}}

The \emph{microxanox} package is for simulating a three functional group
system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
organisms and biogeochemical processes and is based on \citet{Bush2017}
(See \citet{Bush2017} for a detailed discussion of the model). At the
core of the simulations is a set of ordinary differential equations
(specified in the function \texttt{bushplus\_dynamic\_model()}, though
this function need not be directly called). There are functions for
running individual simulations and for running a set of simulations
across, for example, a range of environmental conditions.

To make the simulation run with multiple species per functional group,
we expressed different species characteristics in the elements of
vectors and matrices. We also coded the ordinary differential equations
to include the vectors and matrices, and to use matrix mathematics. In
this way, we made it possible to run simulations with different numbers
of species without having to change the underlying code.

The package functions and code have modular structure so that new
functionality can be easily added. E.g. temporally defined events of any
type could be specified. Further, all parameter values required to run a
simulation are stored in one object. Lastly, the general structure of
the code should make it straightforward to adapt the model to other
similar systems (described in more detail in the Impact section).

\hypertarget{software-architecture}{%
\subsection{Software architecture}\label{software-architecture}}

The framework used when writing this package aimed to maximise
simplicity for the user, and to make it straightforward to reproduce
results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
for an example of how this is used). As such, all the parameters needed
to run a simulation or find a stable state (i.e.~the final state of the
ecosystem) are contained in a single object (which can easily be created
using included functions). This parameter object is given to a function
that runs the simulations and returns the results. The returned results
object is identical to the parameter object but with an additional slot
named \texttt{results}, which contains the results of the run. Thus the
returned results object contains the simulation conditions (parameters)
as well as the results, and can be used to run the simulation again.
This promotes reproducibility and makes incremental changes of
individual parameters and re-running the simulations straightforward.

In the following sections we describe how to use the package to run one
simulation and to find steady states across an environmental gradient.

\hypertarget{running-one-simulation}{%
\subsection{Running one simulation}\label{running-one-simulation}}

A typical simulation would look as shown in Figure
\ref{fig:runsim_example}.

\begin{figure}

{\centering \includegraphics[width=350px]{./figures/simflow}

}

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
\end{figure}

A simulation is run using the \texttt{run\_simulation()} function. In
this function, the ODEs are solved using the function \texttt{ode()} in
the package \emph{deSolve} package \citep{Soetaert2010}. The
\texttt{run\_simulation()} function needs only one argument - an object
as created by the function \texttt{new\_runsim\_parameter()}. The
parameter object returned by \texttt{new\_runsim\_parameter()} contains
among other things the \texttt{strain\_parameter} object, which can be
created by the function \texttt{new\_strain\_parameter()}. For a
detailed description of the parameter objects, their meaning and how
they are created and have values set and changed please see the
\emph{User Guide} which accompanies the package or is available at
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

After the parameter object has been defined, it can be used in the
\texttt{run\_simulation()} function. The function returns an object
which is identical to the parameter object, except of an additional slot
containing the results. This design produces a fully reproducible object
as it can be used as a parameter object to be fed back into the
\texttt{run\_simulation()} function to run the simulation again from the
parameter used to generate the results before.

\hypertarget{finding-stable-states}{%
\subsection{Finding Stable States}\label{finding-stable-states}}

The general approach used to find the stable state of the system with a
specific parameter set is to run the simulation for a long time and
record the final state. When one does this across a range of
environmental conditions, one discovers how the steady state of the
system responds to the environmental conditions. The package contains
functionality for finding steady states that correspond to values of one
environmental driver, namely the value of oxygen diffusivity.

Two methods for finding steady states are implemented. The first runs a
separate simulation for each combination of starting conditions and
oxygen diffusivity (we term this the \emph{Replication method}). This is
the method used in the \citet{Bush2017} study. The second runs two
simulations, one with step-wise and slowly temporally \emph{increasing}
oxygen diffusivity, and the other with step-wise and slowly
\emph{decreasing} oxygen diffusivity. During this temporal environmental
change, the state of the system is recorded just before change to a new
oxygen diffusivity (we term this the \emph{Temporal method}). We
implemented two methods since there is no definitive best method, and in
order to check if results were sensitive to choice of method.

The replication method is implemented in the function
\texttt{run\_replication\_ssfind()} which takes a parameter object as
returned by the function \texttt{new\_replication\_ssfind\_parameter()}
and the number of cores for multithreading the simulation. As the
multithreading uses the package function \texttt{mclapply()} from the
package \texttt{parallel} \citep{RCoreTeam2022}, the multithreading only
works on Linux and Mac. It is planned to move to \texttt{parLapply()}
\citep{RCoreTeam2022} in a future release.

The temporal method is implemented in the function
\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
is planned for a later release, to run these two simulations in
parallel.

For a more detailed walk-through of these two approaches and explanation
please see the
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

\hypertarget{analysing-and-visualising-results}{%
\subsection{Analysing and visualising
results}\label{analysing-and-visualising-results}}

From the results returned, summary measures about how the ecosystem
stable states respond to environmental change can be extracted. The
function \texttt{get\_stability\_measures()} returns quantities such as
the amount of environmental change required for the system to abruptly
change to a different state.

The function \texttt{plot\_dynamics()} plots a single simulation run, as
returned from the \texttt{run\_simulation()} function. This function is
only provided as a convenience function to provide a way to easily see
the results of a simulation run. An example plot resulting from this
function is shown in Figure \ref{fig:plot-dynamics}.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}

}

\caption{Results of a simulation run shown using the function $plot\_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB\_1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB\_1 = phototrophic sulfur bacteria strain 1.}\label{fig:plot-dynamics}
\end{figure}

\hypertarget{impact}{%
\section{Impact}\label{impact}}

The open source implementation and extension of the model used in
\citet{Bush2017} provides the means of reproducing the results published
while at the same time provides the means of doing unique, innovative,
and important investigations of how ecosystems respond to environmental
change, and how biodiversity may modulate this response.

The design of the package code and functionality is with reproducibility
in mind: the combination of all parameters being in a single parameter
object as well as the return of the simulation as a result object which
inherits from the parameter object provides a relatively easy to use
framework to implement reproducible experiments.

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from \citet{Limberger2022} and \citet{Petchey2022}.

\hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
\subsection{Use case 1: Regime shifts during temporal environmental
change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}

The study of \citet{Bush2017} includes simulations of the effect of
oxygen diffusivity (an environmental driver, in the sense that it
affects the ecosystem but is not affected by it) on the ecosystem state
(oxic or anoxic). The \emph{microxonax} package contains functionality
to make a specific temporal pattern of change in the oxygen diffusivity.
As well as allowing individual simulations during which oxygen
diffusivity varies, this functionality forms the basis of the temporal
method for finding stable states.

An example of this functionality is given in the \emph{Partial
Reproduction} vignette, which we briefly show here (Figure
\ref{fig:uc1}). The example is composed of a single simulation, at the
beginning of which the system is in the oxic state with high abundance
of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
eventually, around hour 30'000 the system switches to the anoxic state,
with high abundance of both sulfur bacteria types. The oxygen
diffusivity is then increased and at around hour 38'000 the system
abruptly switches back to the oxic state.

Also visible in the results are thick lines showing abundances of
bacteria when abundances are low. This is due to the implementation of a
function that at regular intervals, increases the abundance to a preset
level. This prevents abundances reducing to very small numbers. The
function that implements this increase abundance can also be made to add
a certain abundance to each strain at regular intervals, thus simulating
immigration in to the system.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/gen_uc1_partrep_temporal_state_switching}

}

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

\hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
\subsection{Use case 2: The extent of hysteresis depends on community
composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}

The package contains a function to extract summary features of ecosystem
responses to environmental change, such as the amount of hysteresis
displayed by the ecosystem. Hysteresis is a key feature of ecosystem
responses to environmental change, because it is related to how
difficult it can be to reserve the effects of environmental change
\citep{Scheffer2001}. The amount of hysteresis is measured as the extent
of the environmental condition (here oxygen diffusitivity) for which
there were two stable states. I.e. it is the extent of the environmental
conditions for which historical conditions play an important role in
determining the current system state (a definition of hysteresis).

Using the package to calculate the extent of hysteresis involves setting
ecosystem and simulation parameters, including parameters for the
finding of stable states across an environmental gradient, running the
stable state finding function, and analysing the results with the
function that calculates extent of hysteresis. The code for this is
provided in the \emph{User Guide}.

The results show that the amount of hysteresis depends greatly on the
combinations of organisms present (Figure \ref{fig:uc2}). For example,
with only the CB (cyanobacteria) present, there was no hysteresis. In
contrast, the presence of both CB and SB (sulfate reducing bacteria) led
to a large amount of hysteresis. (These results are also given in the
\emph{Partial Reproduction} vignette.)

\begin{figure}

{\centering \includegraphics[width=150px]{figures/gen_uc2_user_guide_hysteresis}

}

\caption{The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.}\label{fig:uc2}
\end{figure}

\hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
\subsection{Use case 3: Effects of functional diversity on regime
shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}

As discussed in the Introduction section, the package was motivated by
the question of how biodiversity influences ecosystem responses to
environmental change. Extensive results concerning this question are
given in a separate publication \citet{Limberger2022}. Here we describe
one of the results, which is that having biodiversity in a functional
group can allow state changes to occur that otherwise would not have.
I.e. biodiversity can qualitatively change the state of the ecosystem.

Biodiversity was added to the functional groups using the
\texttt{new\_strain\_parameter()} function to create a parameter set
with multiple species per functional group (albeit all with identical
features) and then to add variability among the species by calling the
\texttt{add\_strain\_var()} function. This function takes an already
existing parameter set and adds the specified about of variation. The
new parameter object is then used as before.

Figure \ref{fig:uc3} shows a simulation with two species (strains) in
each of the three functional groups. The ecosystem starts in the oxic
state, though with relatively high abundance of each functional group.
The strain of SB that is more tolerant to oxygen (SB\_1) initially
decreases in abundance, but then increases, and the other (SB\_2) strain
then becomes abundance and SB\_1 declines. Furthermore, the
cyanobacteria crash in abundance, and the system switches to the anoxic
state. In contrast, if there are two identical strains with tolerance
half way between those in Figure \ref{fig:uc3} the ecosystem remains in
the oxic state.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/uc3_supplement_5_4}

}

\caption{The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.}\label{fig:uc3}
\end{figure}

\hypertarget{adapting-for-other-ecosystems-and-organisms}{%
\subsection{Adapting for other ecosystems and
organisms}\label{adapting-for-other-ecosystems-and-organisms}}

We anticipate that the package can be a useful starting point for
investigating models of other types of ecosystem and how biodiversity in
them affects responses to environmental change. The overall framework of
the package, the purpose of each function, and the objects used for
storing parameters and results could be retained. E.g., all such models
would have parameters that differ among species and need to be described
in an object, and studies will often need to run simulations and sets of
simulations across environmental conditions.

Researchers wanting to model a new ecosystem do not, therefore, have to
start from scratch. This will relieve researchers from needing to make
software design decisions, and rather focus on appropriately
representing their ecosystem and finding the results that interest them.
Nevertheless, adaptation of the code in the package will require a
person / persons that can take a conceptual model of an ecosystem and
then represent that in terms of parameters and rate equations, and that
is relatively proficient in R programming.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

The \emph{microxanox} R package allows the simulation, visualisation,
and analysis of a model of a microbial ecosystem while allowing
variation in the amount of diversity containing in each of the
functional groups of organisms present. It has been used for the
research described in another paper that describes one of the first
investigations of the effects of diversity on ecosystem resilience
\citet{Limberger2022}. In that paper, we show that diversity can have
large and important effects of ecosystem responses, highlighting the
need for models such as ours, with which one can easily manipulate the
amount of biodiversity. The \emph{microxanox} package has also been used
to reproduce the results of the paper that inspired the package
development \citep{Bush2017}.

The package greatly lowers the amount of work required in further
investigations of the specific ecosystem modelled. There has, for
example, been quite limited investigation of how biodiversity influences
the short-term responses of the modelled ecosystem to environmental
change. Likewise, the package could be used to power an investigation of
the effects of biodiversity on the usefulness of early warning signals
of abrupt ecosystem change \citep{Scheffer2009}. In addition this
package could be used as a template for the implementation for
developing models of other types of ecosystems and organism. By doing
so, other models can profit from the overall framework used, and the
reproducibility aspects as well as the flexibility implemented.

\hypertarget{conflict-of-interest}{%
\section{Conflict of Interest}\label{conflict-of-interest}}

The authors declare no known conflicting or competing interests
associated with this publication and there has been no significant
financial support for this work that could have influenced its outcome.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This project was part of SNF Project 310030\_188431. The project was
also supported by the University of Zurich Research Priority Programme
in Global Change and Biodiversity.

\renewcommand\refname{References}
\bibliography{bibliography.bib}

\end{document}