### 2023-06-02 inst/manuscript/manuscript_microxanox.tex
Nearly all done...

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
% 
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
  pdfkeywords={reproducibility, regime shift, steady state, ordinary
differential equations},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial
ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
\author{true \and true}
\date{2023-06-02}

\begin{document}
\maketitle
\begin{abstract}
\(Microxanox\) is an R package to simulate a three functional group
ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
sulfate-reducing bacteria) with four chemical substrates (phosphorus,
oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
differential equations. Simulations can be run individually or over a
parameter range, to find steady states. The model can be implemented
with different numbers of species per functional group. The package is
constructed in such a way that the results contain the input parameter
used, so that a saved results can be loaded again and the simulation be
repeated. Furthermore, the package framework and code should serve as a
useful starting point for making simulation models of other types of
ecosystems.
\end{abstract}

\pagebreak

\hypertarget{required-metadata}{%
\section{Required Metadata}\label{required-metadata}}

\hypertarget{current-code-version}{%
\subsection{Current code version}\label{current-code-version}}

Ancillary data table required for subversion of the codebase.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & v0.9.0 \\
C2 & Permanent links to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide} \\
C9 & Support email for questions &
\href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}};
\href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
\end{longtable}

\hypertarget{motivation-and-significance}{%
\section{Motivation and
significance}\label{motivation-and-significance}}

Mathematical models play a key role in the development of ecosystem
models and the understanding about how and why ecosystems change when
the environment changes (\protect\hyperlink{ref-Binzer2016a}{Binzer et
al., 2016};
\protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021};
\protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}). They are
also important for developing hypotheses to test in empirical studies.
One area of ecology in which models have been influential is the
understanding of their response to a gradual change in an environmental
driver (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). An
environmental driver is an environmental condition that affects an
ecosystem. Still, it is assumed not to be affected by the ecosystem,
such as the rate of nutrient input into a lake.

It is conceivable that an ecosystem state, such as the total biomass of
a particular type of bacteria, may remain unchanged when an
environmental driver changes. It is also possible that the ecosystem
state changes gradually. It is also possible that the ecosystem state
changes abruptly to a new state that is difficult to recover from
(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). This
possibility for abrupt, perhaps catastrophic changes that are difficult
to reverse causes considerable concern
(\protect\hyperlink{ref-Collins2021}{Collins et al., 2021};
\protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021};
\protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).

An example where a gradual change of an environmental variable causes an
abrupt change in the system is the switch from an aerobic (oxygen is
available for metabolism) to an anaerobic (oxygen generally unavailable)
state in a microbial ecosystem. This system has been investigated by
Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}) in a simulation
study of a mathematical model. Three types of microbes occur in the
model: cyanobacteria (CB) dominating the oxic state, and two types of
sulfur bacteria that dominate the anoxic state (sulfate reducing
bacteria (SB) and phototrophic sulfur bacteria (PB)). The model shows
that gradual change in the rate at which oxygen could diffuse into the
ecosystem (termed oxygen diffusitivity) could cause catastrophic changes
in the ecosystem state that would be difficult to reverse.

One feature of the study by Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}) was limited biodiversity.
Specifically, there was no biodiversity within each of the three types
of bacteria. This leaves the question of if and how bacteria's
biodiversity within these types (i.e., functional groups) affects the
ecosystem response to environmental change. This limitation is not
specific to the study of Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}). There are few if any studies
of the effects of biodiversity on abrupt transitions between ecosystem
states.

We decided to fill this research gap by making a simulation study of how
within functional group biodiversity affects ecosystem responses to
environmental change Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
the work and model of Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}). With this goal in mind, we
developed the microxanox package (\protect\hyperlink{ref-Krug2022}{Krug
and Petchey, 2022}). The first stage of development was to write code
from scratch (as there was no available code to start from) and to
confirm that this new implementation would reproduce the previously
published results. The resulting reproduction is available as one of the
package vignettes:
\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
Partial reproduction of Bush et al}.

The second stage was to add functionality that would be necessary to
answer our research question. Most importantly, we made it possible to
have multiple species of bacteria within each of the three functional
groups, for the multiple species to differ in their characteristics, and
to vary the number of species and amount of variability among them. We
also added functionality that allowed: temporally varying environmental
conditions, the addition of random noise to state variables, and
immigration. In addition to the model itself, the package includes some
functions to analyse the results and visualize these to provide a
starting point for customized visualizations based on own requirements.
The basic and additional functionality is described in the package
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

\hypertarget{software-description}{%
\section{Software description}\label{software-description}}

The \emph{microxanox} package is for simulating a three functional group
system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
organisms and biogeochemical processes and is based on Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}) (See Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of
the model). At the core of the simulations is a set of ordinary
differential equations (specified in the function
\texttt{bushplus\_dynamic\_model()}, though this function need not be
directly called). There are functions for running individual simulations
and a set of simulations across, for example, a range of environmental
conditions.

To make the simulation run with multiple species per functional group,
we expressed different species characteristics in the elements of
vectors and matrices. We also coded the ordinary differential equations
to include the vectors and matrices and used matrix mathematics. In this
way, we made it possible to run simulations with different numbers of
species without having to change the underlying code.

The package functions and code have a modular structure so that new
functionality can be easily added. E.g., temporally defined events of
any type could be specified. Further, all parameter values required to
run a simulation are stored in one object. Lastly, the general structure
of the code should make it straightforward to adapt the model to other
similar systems (described in more detail in the Impact section).

\hypertarget{software-architecture}{%
\subsection{Software architecture}\label{software-architecture}}

The framework used when writing this package aimed to maximise
simplicity for the user and to make it straightforward to reproduce
results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey
et al., 2022}) to Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
is used). As such, all the parameters needed to run a simulation or find
a steady state (i.e.~the final state of the ecosystem) are contained in
a single object (which can easily be created using included functions).
This parameter object is given to a function that runs the simulations
and returns the results. The returned results object is identical to the
parameter object but with an additional slot named \texttt{results},
which contains the simulation results. Thus the returned results object
contains the simulation conditions (parameters) as well as the results,
and can be used to run the simulation again. This promotes
reproducibility and makes incremental changes of individual parameters
with a consecutive re-running of the simulations straightforward.

In the following sections we describe how to use the package to run one
simulation and to find steady states across an environmental gradient.

\hypertarget{running-one-simulation}{%
\subsection{Running one simulation}\label{running-one-simulation}}

A typical simulation would look as shown in Figure
\ref{fig:runsim_example}.

\begin{figure}

{\centering \includegraphics[width=350px]{./figures/simflow}

}

\caption{Typical flow of a simulation. Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
\end{figure}

A simulation is run using the \texttt{run\_simulation()} function. In
this function, the ODEs are solved using the function \texttt{ode()} in
the package \emph{deSolve} package
(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}). The
\texttt{run\_simulation()} function needs only one argument - an object
as created by the function \texttt{new\_runsim\_parameter()}. The
parameter object returned by \texttt{new\_runsim\_parameter()} contains,
among other things, the \texttt{strain\_parameter} object, which can be
created by the function \texttt{new\_strain\_parameter()}. For a
detailed description of the parameter objects, their meaning and how
they are created and have values set and changed, please see the
\emph{User Guide} which accompanies the package or is available at
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

After the parameter object has been defined, it can be used in the
\texttt{run\_simulation()} function. The function returns an object
identical to the parameter object, except of an additional slot
containing the results. This design produces a fully reproducible object
as it can be used as a parameter object to be fed back into the
\texttt{run\_simulation()} function to rerun the simulation again from
the parameter used to generate the results before.

\hypertarget{finding-steady-states}{%
\subsection{Finding Steady States}\label{finding-steady-states}}

The general approach to finding the numerical value of a system's steady
state (if it exists) with a specific parameter set is to run the
simulation for a long time and record the final state. When one does
this across a range of environmental conditions, one discovers how the
steady state of the system changes with the environmental conditions.
The package contains functionality for finding steady states that
correspond to values of one environmental driver, namely the value of
oxygen diffusivity.

Two numerical strategies for finding steady states and their sensitivity
to parameters are implemented. Two strategies are implemented in order
to allow comparison of their results. The first method runs a
independent simulation for each combination of initial conditions and
oxygen diffusivity (we term this the \emph{Replication method}). This is
the method used in the Bush et al.
(\protect\hyperlink{ref-Bush2017}{2017}) study and was used to obtain
the results in figures 3 and 4 of that article.

The second method runs two simulations, one with step-wise and slowly
temporally \emph{increasing} oxygen diffusivity, and the other with
step-wise and slowly \emph{decreasing} oxygen diffusivity. (Put another
way, to explore the sensitivity of steady states of the dynamic model
under oxygen diffusivity variation, subset values of this parameter are
numerically evaluated. Two value vectors of oxygen diffusivity are
presented step-wise, gradually increasing and decreasing.) During this
temporal environmental change, the state of the system is recorded just
before change to a new oxygen diffusivity (we term this the
\emph{Temporal method}).

The package does not include methods for a formal analysis of the
stability of the system and users should take care to assess if steady
states are unique, and indeed if a steady state has been achieved.

The replication method is implemented in the function
\texttt{run\_replication\_ssfind()} which takes a parameter object as
returned by the function \texttt{new\_replication\_ssfind\_parameter()}
and the number of cores for multithreading the simulation. As the
multithreading uses the package function \texttt{mclapply()} from the
package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core
Team, 2022}), the multithreading only works on Linux and Mac. It is
planned to move to \texttt{parLapply()}
(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future
release.

The temporal method is implemented in the function
\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
is planned for a later release, to run these two simulations in
parallel.

For a more detailed walk-through of these two approaches and explanation
please see the
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
Guide}.

\hypertarget{analysing-and-visualising-results}{%
\subsection{Analysing and visualising
results}\label{analysing-and-visualising-results}}

From the results returned, summary measures about how the ecosystem
steady states changes with environmental change can be extracted. The
function \texttt{get\_stability\_measures()} returns quantities such as
the amount of environmental change required for the system to abruptly
change to a different state.

The function \texttt{plot\_dynamics()} plots a single simulation run, as
returned from the \texttt{run\_simulation()} function. This function is
only provided as a convenience function to provide a way to easily see
the results of a simulation run. An example plot resulting from this
function is shown in Figure \ref{fig:plot-dynamics}.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}

}

\caption{Results of a simulation run shown using the function $plot\_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB\_1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB\_1 = phototrophic sulfur bacteria strain 1.}\label{fig:plot-dynamics}
\end{figure}

\hypertarget{impact}{%
\section{Impact}\label{impact}}

The open source implementation and extension of the model used in Bush
et al. (\protect\hyperlink{ref-Bush2017}{2017}) provides the means of
reproducing the results published while at the same time provides the
means of doing unique, innovative, and important investigations of how
ecosystems change as the environmental changes, and how biodiversity may
modulate this.

The design of the package code and functionality is with reproducibility
in mind: the combination of all parameters being in a single parameter
object as well as the return of the simulation as a result object which
inherits from the parameter object provides a relatively easy to use
framework to implement reproducible experiments.

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}) and Petchey et al.
(\protect\hyperlink{ref-Petchey2022}{2022}).

\hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
\subsection{Use case 1: Regime shifts during temporal environmental
change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}

The study of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017})
includes simulations of the effect of oxygen diffusivity (an
environmental driver, in the sense that it affects the ecosystem but is
not affected by it) on the ecosystem state (oxic or anoxic). The
\emph{microxonax} package contains functionality to make a specific
temporal pattern of change in the oxygen diffusivity. As well as
allowing individual simulations during which oxygen diffusivity varies,
this functionality forms the basis of the temporal method for finding
steady states.

An example of this functionality is given in the \emph{Partial
Reproduction} vignette, which we briefly show here (Figure
\ref{fig:uc1}). The example is composed of a single simulation, at the
beginning of which the system is in the oxic state with high abundance
of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
eventually, around hour 30'000 the system switches to the anoxic state,
with high abundance of both sulfur bacteria types. The oxygen
diffusivity is then increased and at around hour 38'000 the system
abruptly switches back to the oxic state.

Also visible in the results are thick lines showing abundances of
bacteria when abundances are low. This is due to the implementation of a
function that at regular intervals, increases the abundance to a preset
level. This prevents abundances reducing to very small numbers. The
function that implements this increase abundance can also be made to add
a certain abundance to each strain at regular intervals, thus simulating
immigration in to the system.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/gen_uc1_partrep_temporal_state_switching}

}

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the steady states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

\hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
\subsection{Use case 2: The extent of hysteresis depends on community
composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}

The package contains a function to extract summary features of ecosystem
responses to environmental change, such as the amount of hysteresis
displayed by the ecosystem. Hysteresis is a key feature of ecosystem
responses to environmental change, because it is related to how
difficult it can be to reserve the effects of environmental change
(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). The
amount of hysteresis is measured as the extent of the environmental
condition (here oxygen diffusitivity) for which there were two steady
states. I.e. it is the extent of the environmental conditions for which
historical conditions play an important role in determining the current
system state (a definition of hysteresis).

Using the package to calculate the extent of hysteresis involves setting
ecosystem and simulation parameters, including parameters for the
finding of steady states across an environmental gradient, running the
steady state finding function, and analysing the results with the
function that calculates extent of hysteresis. The code for this is
provided in the \emph{User Guide}.

The results show that the amount of hysteresis depends greatly on the
combinations of organisms present (Figure \ref{fig:uc2}). For example,
with only the CB (cyanobacteria) present, there was no hysteresis. In
contrast, the presence of both CB and SB (sulfate reducing bacteria) led
to a large amount of hysteresis. (These results are also given in the
\emph{Partial Reproduction} vignette.)

\begin{figure}

{\centering \includegraphics[width=150px]{figures/gen_uc2_user_guide_hysteresis}

}

\caption{The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.}\label{fig:uc2}
\end{figure}

\hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
\subsection{Use case 3: Effects of functional diversity on regime
shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}

As discussed in the Introduction section, the package was motivated by
the question of how biodiversity influences ecosystem responses to
environmental change. Extensive results concerning this question are
given in a separate publication Limberger et al.
(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
the results, which is that having biodiversity in a functional group can
allow state changes to occur that otherwise would not have. I.e.
biodiversity can qualitatively change the state of the ecosystem.

Biodiversity was added to the functional groups using the
\texttt{new\_strain\_parameter()} function to create a parameter set
with multiple species per functional group (albeit all with identical
features) and then to add variability among the species by calling the
\texttt{add\_strain\_var()} function. This function takes an already
existing parameter set and adds the specified about of variation. The
new parameter object is then used as before.

Figure \ref{fig:uc3} shows a simulation with two species (strains) in
each of the three functional groups. The ecosystem starts in the oxic
state, though with relatively high abundance of each functional group.
The strain of SB that is more tolerant to oxygen (SB\_1) initially
decreases in abundance, but then increases, and the other (SB\_2) strain
then becomes abundance and SB\_1 declines. Furthermore, the
cyanobacteria crash in abundance, and the system switches to the anoxic
state. In contrast, if there are two identical strains with tolerance
half way between those in Figure \ref{fig:uc3} the ecosystem remains in
the oxic state.

\begin{figure}

{\centering \includegraphics[width=350px]{figures/uc3_supplement_5_4}

}

\caption{The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.}\label{fig:uc3}
\end{figure}

\hypertarget{adapting-for-other-ecosystems-and-organisms}{%
\subsection{Adapting for other ecosystems and
organisms}\label{adapting-for-other-ecosystems-and-organisms}}

We anticipate that the package can be a useful starting point for
investigating models of other types of ecosystem and how biodiversity in
them affects responses to environmental change. The overall framework of
the package, the purpose of each function, and the objects used for
storing parameters and results could be retained. E.g., all such models
would have parameters that differ among species and need to be described
in an object, and studies will often need to run simulations and sets of
simulations across environmental conditions.

Researchers wanting to model a new ecosystem do not, therefore, have to
start from scratch. This will relieve researchers from needing to make
software design decisions, and rather focus on appropriately
representing their ecosystem and finding the results that interest them.
Nevertheless, adaptation of the code in the package will require a
person / persons that can take a conceptual model of an ecosystem and
then represent that in terms of parameters and rate equations, and that
is relatively proficient in R programming.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

The \emph{microxanox} R package allows the simulation, visualisation,
and analysis of a model of a microbial ecosystem while allowing
variation in the amount of diversity containing in each of the
functional groups of organisms present. It has been used for the
research described in another paper that describes one of the first
investigations of the effects of diversity on ecosystem resilience
Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2022}). In that
paper, we show that diversity can have large and important effects of
ecosystem responses, highlighting the need for models such as ours, with
which one can easily manipulate the amount of biodiversity. The
\emph{microxanox} package has also been used to reproduce the results of
the paper that inspired the package development
(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).

The package greatly lowers the amount of work required in further
investigations of the specific ecosystem modelled. There has, for
example, been quite limited investigation of how biodiversity influences
the short-term responses of the modelled ecosystem to environmental
change. Likewise, the package could be used to power an investigation of
the effects of biodiversity on the usefulness of early warning signals
of abrupt ecosystem change
(\protect\hyperlink{ref-Scheffer2009}{Scheffer et al., 2009}). In
addition this package could be used as a template for the implementation
for developing models of other types of ecosystems and organism. By
doing so, other models can profit from the overall framework used, and
the reproducibility aspects as well as the flexibility implemented.

\hypertarget{conflict-of-interest}{%
\section{Conflict of Interest}\label{conflict-of-interest}}

The authors declare no known conflicting or competing interests
associated with this publication and there has been no significant
financial support for this work that could have influenced its outcome.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This project was part of SNF Project 310030\_188431. The project was
also supported by the University of Zurich Research Priority Programme
in Global Change and Biodiversity.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Binzer2016a}{}}%
Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016. Interactive effects
of warming, eutrophication and size structure: Impacts on biodiversity
and food-web structure. Global Change Biology 22, 220--227.
doi:\href{https://doi.org/10.1111/gcb.13086}{10.1111/gcb.13086}

\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
2017. Oxic-anoxic regime shifts mediated by feedbacks between
biogeochemical processes and microbial community dynamics. Nature
Communications 8, 789.
doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}

\leavevmode\vadjust pre{\hypertarget{ref-Chaparro-Pedraza2021}{}}%
Chaparro-Pedraza, P.C., 2021. Fast environmental change and
eco-evolutionary feedbacks can drive regime shifts in ecosystems before
tipping points are crossed. Proceedings of the Royal Society B:
Biological Sciences 288, 20211192.
doi:\href{https://doi.org/10.1098/rspb.2021.1192}{10.1098/rspb.2021.1192}

\leavevmode\vadjust pre{\hypertarget{ref-Collins2021}{}}%
Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P.,
Ratajczak, Z., 2021. Fire frequency, state change and hysteresis in
tallgrass prairie. Ecology Letters 24, 636--647.
doi:\href{https://doi.org/10.1111/ele.13676}{10.1111/ele.13676}

\leavevmode\vadjust pre{\hypertarget{ref-Krug2022}{}}%
Krug, R.M., Petchey, O.L., 2022. Microxanox: {Model} to simulate three
functional group system.
doi:\href{https://doi.org/10.5281/ZENODO.6624125}{10.5281/ZENODO.6624125}

\leavevmode\vadjust pre{\hypertarget{ref-Limberger2022}{}}%
Limberger, R., Daugaard, U., Gupta, A., Krug, R., Lemmen, K., Moorsel,
S. van, Suleiman, M., Zuppinger-Dingley, D., Petchey, O., 2022.
Functional diversity can facilitate the collapse of an undesirable
ecosystem state. Authorea Preprint.
doi:\href{https://doi.org/10.22541/au.165513066.66785786/v1}{10.22541/au.165513066.66785786/v1}

\leavevmode\vadjust pre{\hypertarget{ref-Northrop2021}{}}%
Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J.,
2021. Clockwise and counterclockwise hysteresis characterize state
changes in the same aquatic ecosystem. Ecology Letters 24, 94--101.
doi:\href{https://doi.org/10.1111/ele.13625}{10.1111/ele.13625}

\leavevmode\vadjust pre{\hypertarget{ref-Petchey2022}{}}%
Petchey, O.L., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., van
Moorsel, Sofia J., Suleiman, M., Zuppinger-Dingley, Debra, Limberger,
R., 2022. Code package for: "{Functional} diversity can facilitate the
collapse of an undesirable ecosystem state".
doi:\href{https://doi.org/10.5281/ZENODO.6334544}{10.5281/ZENODO.6334544}

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
R Core Team, 2022. R: {A} language and environment for statistical
computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.

\leavevmode\vadjust pre{\hypertarget{ref-Scheffer2009}{}}%
Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R.,
Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
Early-warning signals for critical transitions. Nature 461, 53--59.
doi:\href{https://doi.org/10.1038/nature08227}{10.1038/nature08227}

\leavevmode\vadjust pre{\hypertarget{ref-Scheffer2001}{}}%
Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001.
Catastrophic shifts in ecosystems. Nature 413, 591--596.
doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}

\leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
1--25.
doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}

\leavevmode\vadjust pre{\hypertarget{ref-Vandermeer2019}{}}%
Vandermeer, J., Perfecto, I., 2019. Hysteresis and critical transitions
in a coffee agroecosystem. Proceedings of the National Academy of
Sciences 116, 15074--15079.
doi:\href{https://doi.org/10.1073/pnas.1902773116}{10.1073/pnas.1902773116}

\leavevmode\vadjust pre{\hypertarget{ref-Vasseur2005}{}}%
Vasseur, D.A., McCann, K.S., 2005. A {Mechanistic Approach} for
{Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}. The
American Naturalist 166, 184--198.
doi:\href{https://doi.org/10.1086/431285}{10.1086/431285}

\end{CSLReferences}

\end{document}